---
title: "W6 assignment"
author: "Bianka Sz."
date: "10/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## W6 
1. Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.

```{r }
#loading packages, data, putting data into dataframe
library(pacman)
p_load(tidyverse, gapminder)
df <- gapminder

#creating function
GDP_calc <- function(dat, year=NULL, country=NULL) {
  if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
    dat <- dat[dat$country %in% country,]
  }
  gdp_nation <- dat$pop * dat$gdpPercap

  new <- cbind(dat, gdpNation=gdp_nation)
  return(new)
}

#defining the years I want
years <- c(1967, 1977, 1987, 1997, 2007, 2017)

#applying the function to the dataframe and putting the output into a new dataframe
Dk <- GDP_calc(df,country = "Denmark", year= years)
#Viewing the new dataframe, NB: the year 2017 is not added, because it was not included in the original dataframe
View(Dk)
```

2. Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.
```{r }
#finding the countries, whose name start with a 'B' and assigning them into a list
B_countries <- grep("^B", unique(gapminder$country), value=TRUE)
#setting the tresholds
lowerThreshold <- 50
upperThreshold <- 70
 
for (iCountry in B_countries) {
    tmp <- mean(df$lifeExp[df$country == iCountry])
    
    if(tmp < lowerThreshold){
        cat("Average Life Expectancy in", iCountry, "is less than", lowerThreshold, "\n")
    }
    else if(tmp > lowerThreshold && tmp < upperThreshold){
        cat("Average Life Expectancy in", iCountry, "is between", lowerThreshold, "and", upperThreshold, "\n")
    }
    else{
        cat("Average Life Expectancy in", iCountry, "is greater than", upperThreshold, "\n")
    }
    rm(tmp)
}
```
Optional: Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘M’ and graphs life expectancy against time (using plot() function) as a line graph if the mean life expectancy is under 50 years.
Hint: If you are struggling with the gapminder tibble format, consider converting it into a dataframe (from tibble), either by downloading it from the internet and loading it via read.csv (not read_csv), and/or using as.data.frame() conversion function and then appropriately subsetting.
