---
title: "W5 assignment"
author: "Bianka Sz."
date: "3/7/2021"
output: html_document
---
## Task 2: Where in the landscape are the mounds located?

The location of settlements is usually easy to predict as humans need safety and easy access to water and food resources. These range from nearby arable soil to pasturage to trading centers. Choices regarding the mortuary realm are much harder to establish as they are guided by the social norms of the living rather than the natural environment. Some environmental conditions, however, play a role, such as intervisibility, which can be an important factor for societies that use monuments to the dead for territorial signalling. Before such specific analysis, it is, however, a good idea to get a general sense of where in the landscape are mounds located.

In order to produce a formal assessment of mound location, you can start by using most common aspects of topography, such as elevation, slope, aspect, and water proximity. Choose one or more of these variables. Calculate the distribution of classes of each environmental variable through the entire region (defining, en route, what you consider to be a “region”?) by extracting values from the digital elevation model and binning them in several classes. Then, calculate site frequencies within classes of the environmental variable by sampling mound locations in the raster and evaluate whether the observed pattern could be a product of the distribution of environmental categories themselves.

A example workflow with elevations could look like this:

-extract elevation values from the entire landscape and bin them in elevation categories (e.g. 400-500m, 500-600m, 600-700m, etc.). Consider: what defines a landscape for the purpose of this study? You may wish to crop the Aster to a smaller but representative study area but make sure you justify your selection
-extract elevation values from observed mound locations and review their distribution in the same bins
-calculate the expected number of mounds per bin if mounds were located through the landscape randomly
-compare the expected number with the observed one

```{r }
# loading in libraries
library(sf)
library(raster)
library(lattice)
library(latticeExtra)
library(RColorBrewer)
library(rasterVis)
library(rgdal)
library(rgl)
library(tmap)
library(ggplot2)
library(tidyverse)

# reading in the data
aster <- raster("data/Aster.tif")
mounds <- st_read("data/KAZ_mounds.shp")
```

## Preparing maps

```{r }
# checking crs
st_crs(mounds)
crs(aster) 
#they are both WGS84/UTM 35N

#plotting aster elevations to visually inspect them
plot(aster)

#we have to remove negative values
rclmat <- cbind(-10000,-1, NA)
aster <- reclassify(aster, rcl = rclmat)

#now plottingthe reclassified aster to see improvement 
plot(aster) #looks way better

#mapping the mounds on top of the aster 
original_map <- tm_shape(aster)+
  tm_raster()+
  tm_shape(mounds)+
  tm_markers(size = 0.1)+
  tm_layout(main.title = "Original map")

#cropping the aster object to only contain the mounds 
aster_crop <- crop(aster, mounds)

#mapping the mounds on top of the cropped aster
cropped_map <- tm_shape(aster_crop)+
  tm_raster()+
  tm_shape(mounds)+
  tm_markers(size = 0.1)+
  tm_layout(main.title = "Cropped map")

#plotting both of the maps
tmap_arrange(original_map,cropped_map, asp = NA)
```

## Elevation

```{r }
#extracting elevation values from the cropped aster
aster_crop_values <- values(aster_crop)

# checking max and min values
max(aster_crop_values) #1302
min(aster_crop_values) #265

#defining elevation breaks according to the min and max values
breaks = c(200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200, 1300, 1400)

#plotting the histogram of the elevation values in the cropped aster
hist(aster_crop_values, breaks = breaks)

#calculating the elevation frequencies
elevbins <- cut(values(aster_crop), breaks = breaks, dig.lab = 5) 
elev_freq <- data.frame(table(elevbins))
elev_freq

#extracting the elevation for the mounds
mounds$elevation <- raster::extract(aster_crop, mounds)

#plotting mounds histogram
hist(mounds$elevation, breaks = breaks)

#calculating elevation frequency for mounds
elevbins <- cut(mounds$elevation, breaks = breaks, dig.lab = 5)
mounds_freq <- data.frame(table(elevbins))
mounds_freq
```
My conclusion is: The burial mounds are generally situated rather low. Specifically, between the elevation of 300-700 m above sea level, with a strong preference for 300-500 m asl.
